### 2019-12-16 paper/conclusion.tex

\section{Conclusion} \label{Conclusion}

\section{Future work}

>>>>

\section{Conclusion} \label{Conclusion}
In the current work, we introduced design of reactivity control system for the 
Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR). The 
design has been evaluated at the startup for the full-core geometry using 
SERPENT-2 Monte-Carlo code. We considered three various startup composition of 
the fuel salt: (1) $^{232}$Th as fertile and $^{233}$U as fissile material; 
(2) $^{232}$Th and plutonium, extracted from \gls{LWR} \gls{SNF}; (3) 
$^{232}$Th and transuranic vector, extracted from \gls{LWR} \gls{SNF}.

The excess reactivity $\rho_e$ is calculated at zero burnup when all CRs are 
fully withdrawn. $\rho_e$ for $^{233}$U, reactor-grade Pu, and 
TRU are $1.65\pm0.04$ $\$$, $4.11\pm0.02$ $\$$, and $15.38\pm0.04$ $\$$, 
respectively.

Five different absorber materials are considered based on their neutronics and 
safety performance: B$_4$C-90 (boron is enriched to 90\% $^{10}$B), HfB$_2$, 
HfH$_{1.62}$, Eu$_2$O$_3$, and Gd$_2$O$_3$. Enriched B$_4$C-90 has the largest 
absorption ability, while Gd$_2$O$_3$ has the lowest absorption compared with 
the other absorber materials in this study. Both CSD and SSD clusters are 
separately able to shut down the reactor initially loaded with $^{233}$U and 
reactor-grade Pu regardless of the absorber material type. However, only SSD 
clusters made of B$_4$C-90 is able to shut down the SD-TMSR initially loaded 
with transuranium vector from \gls{LWR}. The reason for this is much larger 
absorption cross section of $^{10}$B in the relatively soft neutron energy 
spectrum of the SD-TMSR core started with TRU.

In case of malfunction of the other CR clusters (e.g., stuck in the upper 
position), the outer ring of the CSD failures to counteract the excess 
reactivity of the core initially loaded with reactor-grade Pu and transuranic 
(TRU) elements. However, the worth of the outer ring of the CSD is sufficient 
to compensate the excess reactivity for the core refueled by $^{233}$U.

All absorber materials provide an adequate shutdown margin for the SD-TMSR 
core that initially loaded with $^{233}$U and reactor-grade Pu. However, the 
shutdown margins for TRU case are negative or slightly positive (in B$_4$C-90 
case), this makes the SSD clusters ineffective to shut down the reactor loaded 
with TRU.

The amplification factor (A$_{CRi}$) results show that the CSD, SSD, CSD inner ring, and SSD1 are slightly amplified due to the anti-shadowing effects. The strongest anti-shadowing effect has observed in SSD4 and CSD9 clusters that are located at the boundary between the core zones with different moderator-to-fuel ratio.

The insertion of CRs deforms the radial flux shape in certain positions, i.e., around CRs positions. This shifts the neutron flux from the core center towards the periphery.

The integral and differential control rod worth are calculated for three 
different systems: all control rods, CSD, and SSD systems. The results show 
that all absorber materials have almost the same integral rod worth in the 
upper half of the core. Further insertion of the control rods shows the unique 
absorption characteristics of each material.

\section{Future work}
In Part II of this research effort we will evaluate effect of fuel salt 
burnup on the CRW. The depleted fuel composition will be obtained using 
SERPENT-2 online reprocessing subroutine \cite{aufiero2013extended} and 
batch-wise tool SaltProc \cite{rykhlevskii_arfc/saltproc_2018, 
rykhlevskii_milestone_2019}. Additionally, the authors indent to investigate 
kinetic parameters (effective delayed neutron fraction $\beta_{eff}$ and 
effective delayed neutron precursor decay constant $\lambda_{eff}$) evolution 
during the SD-TMSR operation. These parameters are crucial for accident 
transient analysis, which will be performed using Moltres 
\cite{lindsay_introduction_2018}, multi-physics code for liquid-fueled MSR 
simulation with taking into account the neutron precursors drift.

### 2019-12-16 paper/methods.tex

The Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) design was proposed by the Chinese Academy of Sciences in 2011 
\cite{li_optimization_2018,jiang2012advanced,li2015analysis,li2017model}. The \gls{SD-TMSR} is a graphite-moderated molten salt reactor with a thermal power of 2,250 MW$_{th}$. The 
design of the \gls{SD-TMSR} is inspired by the \gls{MSBR} 
\cite{robertson_conceptual_1971} and the \gls{TMSR} \cite{nuttin2005potential} after modifying the geometry to 
control the positive moderator temperature coefficient in the MSBR. The \gls{SD-TMSR} core 
geometry was described in detail by Li \emph{et al.} and Ashraf \emph{et al.}
\cite{li_optimization_2018,ashraf2019whole_core}. The active zone of the \gls{SD-TMSR} is a right cylinder divided into the inner zone (486 fuel tubes) and the outer zone (522 fuel tubes) to enhance breeding performance.

In this study, the fuel salt composition is 70LiF - 17.5BeF$_2$ - 12.5(HM)F$_N$ mole\%, where HM is the heavy metal (i.e. $^{232}$Th and fissile materials), and $N$ is dependent on the chosen fissile material and the thermochemical state of the liquid fuel salt. Three different types of initial fissile materials are considered: (1) $^{233}$U \cite{ashraf2019whole_core}, (2) reactor-grade Pu \cite{marka1993explosive}, and (3) transuranic (TRU) elements from \gls{LWR} \gls{SNF} \cite{de2000scenarios}.
The density and volume of the fuel salt are 3.3 g/cm$^{3}$ and 52.9 m$^3$, respectively.
The liquid fuel salt circulates continuously through the fuel tubes that pierce the graphite hexagonal 
prisms. The active zone is surrounded by axial and radial graphite reflectors to minimize the neutron leakage. Finally, the \gls{SD-TMSR} pressure vessel holds all reactor components and is made of a Hastelloy N alloy. The main 
characteristics of the \gls{SD-TMSR} are summarized in Table~\ref{tab:table1}.

@@

The reactivity of the SD-TMSR core is controlled through two systems of control assemblies:
\begin{enumerate}
\item The Control Safety Devices (CSD), and
\item The Shutdown Safety Devices (SSD).
\end{enumerate}
The CSD system is designed for reactivity control during normal operation and the SSD system is designed for an emergency reactor shutdown.
In the present work, five different absorber materials are considered based on their neutronics and safety performance:
\begin{enumerate}
\item B$_4$C with boron enriched to 90\% $^{10}$B isotope,
\item hafnium diboride (HfB$_2$),
\item hafnium hydride (HfH$_{1.62}$),
\item gadolinium oxide (Gd$_2$O$_3$), and
\item europium oxide (Eu$_2$O$_3$).
\end{enumerate}
Boron carbide (B$_4$C) is widely used in control rods fabrication. Enriching B$_4$C increases the fraction of highly neutron absorber isotope (i.e. $^{10}$B) and helps to reach the necessary absorptivity. However, issues related to helium gas release through (n,$\alpha$) reactions of $^{10}$B, swelling, melting risk, and high loss of the reactivity worth limit the B$_4$C lifespan \cite{guo2019optimized}. Hafnium and rare earth elements oxides absorb neutrons through (n,$\gamma$) reactions and this limits gas emission.

The control rod is a cylinder with a radius of 0.75 cm and a height of 520 cm. The absorber material is surrounded by a 0.25-cm-thick clad made of AIM1 (15Cr-15Ni) steel alloy \cite{SERAN2017285} and the rod follower is made of SiC structural material (see Figure~\ref{fig:cr}). A small gap between the cladding and rod follower is considered to facilitate the control rod movement.

Since the total quantity and distribution of the control assemblies in the SD-TMSR are unknown, we proposed an original distribution as a starting point of this analysis. We added clusters consist of four control rods to specific graphite hexagonal prisms (elements) in the SD-TMSR core. Every four control rods (cluster) can move together as one group. Figure~\ref{fig:graphite_elemen} and~\ref{fig:graphite_elemen1} demonstrate the plan and axial view of the graphite element with the control rods. The total numbers of graphite elements with control rods are 25: 16 CSD and 9 SSD.

\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in} 
	\includegraphics[width=\textwidth]{cr.png}
	\caption{Cross section of the B$_4$C control rod.}
	\label{fig:cr}
\end{figure}

\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{graphite_element.png}
	\caption{Graphite element with the four control rods (cluster) located at the same distance from the fuel channel.}
	\label{fig:graphite_elemen}
\end{figure}

\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{graphite_element1.png}
	\caption{Axial view of graphite element with control rods.}
	\label{fig:graphite_elemen1}
\end{figure}

Figure~\ref{fig:core_25} illustrates the numbering scheme of control rods clusters in the SD-TMSR core.
The CSD1-16 clusters are represented as yellow color and distributed as two rings: inner and outer ring (peripheral ring). The inner ring includes CSD from 1 to 6, while the outer ring includes CSD from 7 to 16. Red color stands for SSD1-9 clusters.
We distributed the graphite elements with control rods clusters homogeneously in the inner core of the SD-TMSR, where the moderator-to-fuel ratio is high.
The selected core segment in the above corner of the Figure~\ref{fig:core_25} shows that both CSD and SSD clusters consist of four control rods located at the same distance from the fuel channel.

\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{core_25.png}
	\caption{Distribution of the graphite elements with control rods in the SD-TMSR core.}
	\label{fig:core_25}
\end{figure}

\section{Methodology and tools} \label{Methodology-and-tools}
\subsection{Control rod design evaluation}

In this work, SERPENT-2 version 2.1.31 \cite{leppanen2014serpent} is used to perform steady state calculations for the full-core of the Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) with control rod design. We adopted the ENDF-VII.0 cross section library for all calculations in the present work. The results demonstrate whole-core runs of $12.5\times 10^6$ neutron histories per depletion step. The statistical error in $k_{eff}$ from SERPENT-2 output is equal to $\pm$ $25$ $pcm$.

The initial calculation state of the SD-TMSR is identified by normal operation conditions (see Table~\ref{tab:table1}) and fully withdrawn control rods clusters. In this case, the control rods are located above the upper plenum as shown in Figure~\ref{fig:core_26}. To validate the proposed control rods system we adopted the same operation conditions (as in the initial calculation state) and changed the position of the control rod clusters along $Z$ direction. The main calculated parameters including reactivity, control rod worth, and interference effects (shadowing effects) will describe in the following parts.

@@

Interference between control rods (CRs) or shadowing effects occurs when one (or more) control rod affects the reactivity worth of another control rod in the surroundings. Thus, anti-shadowing is observed when the combined rod worth is greater than the sum of the individual worths, however, the shadowing effect appears when the combined rod worth is less than the sum of the individual worths.
The core height-to-diameter ratio (H/D) and the three-dimensional configuration of the control rods affect the degree of the interference between CRs \cite{girardin2007control}.

@@

The integral CRW is the total reactivity change due to control rod insertion or withdrawal. However, the differential CRW is the amount of reactivity inserted per unit of withdrawal [$\$/cm$]. We change the position of CRs clusters parallel to the z-axis from top to the bottom of the core. Equation~\ref{Equ:4} is used to calculate the integral CRW [\$], where $k_{i}$ and $k_{i-1}$ are the effective multiplication factor after and before CR insertion to i$^{th}$ step. $\beta_{i}$ is the effective fraction of delayed neutrons at i$^{th}$ step. N is the number of steps.

>>>>

The Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) design was proposed by the Chinese Academy of Sciences in 2011 
\cite{li_optimization_2018,jiang2012advanced,li2015analysis,li2017model}. The 
\gls{SD-TMSR} is a graphite-moderated molten salt reactor with a thermal power 
of 2,250 MW$_{th}$. The design of the \gls{SD-TMSR} is inspired by the 
\gls{MSBR} \cite{robertson_conceptual_1971} and the \gls{TMSR} 
\cite{nuttin2005potential} after modifying the geometry to control the 
positive moderator temperature coefficient in the MSBR. The \gls{SD-TMSR} core 
geometry was described in detail by Li \emph{et al.} and Ashraf \emph{et al.}
\cite{li_optimization_2018,ashraf2019whole_core}. The core of the  
\gls{SD-TMSR} is a right cylinder divided into the inner zone (486 fuel tubes) 
and the outer zone (522 fuel tubes) to enhance breeding performance.

In this study, the fuel salt composition is 70LiF - 17.5BeF$_2$ - 
12.5(HM)F$_N$ mole\%, where HM is the heavy metal (i.e. $^{232}$Th and fissile 
materials), and $N$ depends on the chosen fissile material and the  
thermochemical state of the liquid fuel salt. Three different types of initial 
fissile materials are considered: (1) $^{233}$U \cite{ashraf2019whole_core}, 
(2) reactor-grade Pu \cite{marka1993explosive}, and (3) transuranic (TRU) 
elements from \gls{LWR} \gls{SNF} \cite{de2000scenarios}.
The density and volume of the fuel salt are 3.3 g/cm$^{3}$ and 52.9 m$^3$, 
respectively. The liquid fuel salt circulates continuously through the fuel 
tubes that pierce the graphite hexagonal prisms. The core is surrounded by 
axial and radial graphite reflectors to minimize the neutron leakage. Finally, 
the \gls{SD-TMSR} pressure vessel holds all reactor components and is made of 
a Hastelloy N alloy. The main characteristics of the \gls{SD-TMSR} are 
summarized in Table~\ref{tab:table1}.

@@

The reactivity of the SD-TMSR core is controlled by two clusters of control 
rods:
\begin{enumerate}
\item Control Safety Devices (CSD);
\item Shutdown Safety Devices (SSD).
\end{enumerate}
The CSD system is designed for reactivity control during normal operation and the SSD system is designed for an emergency reactor shutdown.
In the present work, five different absorber materials are considered based on their neutronics and safety performance:
\begin{enumerate}
\item B$_4$C with boron enriched to 90\% $^{10}$B;
\item hafnium diboride (HfB$_2$);
\item hafnium hydride (HfH$_{1.62}$);
\item gadolinium oxide (Gd$_2$O$_3$);
\item europium oxide (Eu$_2$O$_3$).
\end{enumerate}
Boron carbide (B$_4$C) is widely used in control rods fabrication. Enriching 
PB$_4$C increases the fraction of highly neutron absorber isotope (i.e. 
$^{10}$B) and helps to reach the necessary absorptivity. However, issues 
related to helium gas release through (n,$\alpha$) reactions of $^{10}$B, 
swelling, melting risk, and high loss of the reactivity worth limit the B$_4$C 
lifespan \cite{guo2019optimized}. Hafnium and rare earth elements oxides 
absorb neutrons through (n,$\gamma$) reactions and this limits gas emission.

The control rod is a cylinder with a radius of 0.75 cm and a height of 520 cm. 
The absorber material is surrounded by a 0.25-cm-thick cladding made of AIM1 
(15Cr-15Ni) steel alloy \cite{SERAN2017285} and the rod follower is made of 
SiC structural material (see Figure~\ref{fig:cr}). A small gap between the 
cladding and rod follower is considered to facilitate the control rod movement.

Since the total quantity and distribution of the control assemblies in the SD-TMSR are unknown, we proposed an original distribution as a starting point of this analysis. We added clusters consist of four control rods to specific graphite hexagonal prisms (elements) in the SD-TMSR core. Every four control rods (cluster) can move together as one group. Figure~\ref{fig:graphite_elemen} and~\ref{fig:graphite_elemen1} demonstrate the plan and axial view of the graphite element with the control rods. The total numbers of graphite elements with control rods are 25: 16 CSD and 9 SSD.
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in} 
	\includegraphics[width=\textwidth]{cr.png}
	\caption{Cross section of the B$_4$C control rod.}
	\label{fig:cr}
\end{figure}
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{graphite_element.png}
	\caption{$XY$ section of graphite element with the four control rods 
	(cluster) located at the same distance from the fuel channel.}
	\label{fig:graphite_elemen}
\end{figure}
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{graphite_element1.png}
	\caption{$XZ$ section of graphite element with control rods.}
	\label{fig:graphite_elemen1}
\end{figure}

Figure~\ref{fig:core_25} illustrates the numbering scheme of control rods 
clusters in the SD-TMSR core.
The CSD1-16 clusters are represented as yellow color and distributed as two rings: inner and outer ring (peripheral ring). The inner ring includes CSD from 1 to 6, while the outer ring includes CSD from 7 to 16. Red color stands for SSD1-9 clusters.
We distributed the graphite elements with control rods clusters uniformly in 
the inner core of the SD-TMSR, where the moderator-to-fuel ratio is high.
The selected core segment at the upper left corner of the 
Figure~\ref{fig:core_25} shows that both CSD and SSD clusters consist of four 
control rods located at the same distance from the fuel channel center.
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{core_25.png}
	\caption{Distribution of the graphite elements with control rods in the SD-TMSR core.}
	\label{fig:core_25}
\end{figure}

\section{Methodology and tools} \label{Methodology-and-tools}
\subsection{Control rod design evaluation}
In this work, SERPENT-2 version 2.1.31 \cite{leppanen2014serpent} is used to 
perform steady state calculations for the full-core of the SD-TMSR with 
suggested control rod design. We adopted the ENDF-VII.0 cross section library 
for all calculations in the present work. The results demonstrate whole-core 
runs of $12.5\times 10^6$ neutron histories per depletion step. The 
statistical error in $k_{eff}$ from SERPENT-2 output is equal to $\pm$ $25$ 
$pcm$.

The initial calculation state of the SD-TMSR is identified by normal operation 
conditions (see Table~\ref{tab:table1}) and fully withdrawn control rods 
clusters. In this case, the control rods are located above the upper plenum as 
shown in Figure~\ref{fig:core_26}. To validate the proposed control rods 
system we adopted the same operation conditions (as in the initial calculation 
state) and changed the position of the control rod clusters along $Z$ 
direction. The main calculated parameters including reactivity, control rod 
worth, and interference effects (shadowing effects) are described below.

@@

Interference between control rods (CRs) or shadowing effects occurs when one 
(or more) control rod affects the reactivity worth of another control rod in 
the surroundings. Thus, anti-shadowing is observed when the combined rod worth 
is greater than the sum of the individual worths. Oppositely, the shadowing 
effect appears when the combined rod worth is less than the sum of the 
individual worths. The core height-to-diameter ratio (H/D) and the 
three-dimensional configuration of the control rods affect the degree of the 
interference between CRs \cite{girardin2007control}.

@@

The integral CRW is the total reactivity change due to the control rod full 
insertion or withdrawal. In contrast, the differential CRW is the amount of 
reactivity inserted per unit of withdrawal [$\$/cm$]. To calculate those 
parameters, we varied the position of CRs clusters from fully withdrawn to 
fully inserted. Equation~\ref{Equ:4} is used to calculate the integral CRW 
[\$], where $k_{i}$ and $k_{i-1}$ are the effective multiplication factor 
after and before CR insertion to i$^{th}$ step. $\beta_{i}$ is the effective 
fraction of delayed neutrons at i$^{th}$ step. N is the number of steps.
